---
title: "NFL Position Value Analysis"
author: "Ibrahim Taher"
date: "September 9, 2018"
output: pdf_document
---

```{r}
library(tidyverse)
library(cluster)
setwd('C:/Users/ibiyt/Desktop/GitHub/fantasy-football/HistoricalFantasyData')
```

```{r}
data<- lapply(list.files(),read_csv)
year=2010

addYear <- function(tbl) {
  tbl$YEAR = rep(year,nrow(tbl))
  year <<- year + 1
  tbl
}

data<-lapply(data,addYear)

dataMerged <- bind_rows(data)
dataMerged <- dataMerged %>%
  mutate(TEAM = if_else(TEAM=='STL','LAR',TEAM)) %>%
  mutate(TEAM = if_else(TEAM=='SDG','LAC',TEAM))
dataMerged
```

```{r}
dataMerged %>% 
  filter(PosRank<=30) %>%
  select(TEAM,POS,PPR,YEAR) %>%
  filter(TEAM!='2TM') %>%
  group_by(TEAM,POS,YEAR) %>%
  mutate(TEAM_RANK = row_number()) %>%
  mutate(TEAM_POS_RANK = paste(POS,TEAM_RANK,sep = '_')) %>%
  arrange(YEAR,TEAM,POS,TEAM_RANK) %>%
  ggplot() + geom_col(mapping = aes(x = YEAR, y = PPR,fill=TEAM_POS_RANK)) + facet_wrap(~TEAM)
```

```{r}
dataMerged %>% 
  filter(PosRank<=30) %>%
  select(TEAM,POS,PPR,YEAR) %>%
  filter(TEAM!='2TM') %>%
  group_by(TEAM,POS,YEAR) %>%
  mutate(TEAM_RANK = row_number()) %>%
  arrange(TEAM,POS,YEAR,PPR) %>%
  ungroup()%>%
  mutate(TEAM_POS_RANK = paste(POS,TEAM_RANK,sep = '_')) %>%
  ggplot() + geom_col(mapping = aes(x = TEAM, y = PPR,fill=TEAM_POS_RANK)) + facet_wrap(~YEAR,scales = 'free_y') + coord_flip()
```

An interesting idea: what if we took the QB1, RB1, RB2, WR1, WR2, TE1 of each team and clustered them. 
Would ideally create clusters based on PPR performance of each player and we'd have tiers. Or should we just do it by total. Let's do both.

```{r}
dataMerged %>%
  select(TEAM,POS,PPR,YEAR) %>%
  group_by(TEAM,POS,YEAR) %>%
  mutate(TEAM_RANK = row_number()) %>%
  filter(TEAM_RANK<=2 & TEAM!='3TM' & TEAM!='4TM' & YEAR>=2014) %>%
  arrange(TEAM,POS,YEAR,PPR) %>% 
  filter(TEAM!='2TM') %>%
  unite(POS_TEAM_RANK,POS,TEAM_RANK,sep = '_') %>%
  select(TEAM,POS_TEAM_RANK,PPR,YEAR) %>%
  filter(!POS_TEAM_RANK %in% c('QB_2','NA_1','NA_2','RB_2','TE_2')) %>%
  ggplot() + geom_line(mapping = aes(x=YEAR,y=PPR,color=POS_TEAM_RANK)) + facet_wrap(~TEAM)
```

While messing around with the number of years shown, what I have discovered is
that there isn't any realy defined pattern because those people who fill those
positions can regress or advance really quickly. We do see certain trends,
like PIT WR_1 and MIN WR_1 both just getting better with time. Detroits been
slightly regressing, though not by much.

```{r}
dataMerged %>%
  select(TEAM,PPR,YEAR) %>%
  group_by(TEAM,YEAR) %>%
  mutate(TEAM_RANK = row_number()) %>%
  filter(TEAM_RANK<=2 & TEAM!='2TM' &TEAM!='3TM' & TEAM!='4TM') %>%
  group_by(TEAM,YEAR) %>%
  summarise(PPR=sum(PPR)) %>%
  ggplot() + geom_line(mapping = aes(x=YEAR,y=PPR,color=TEAM))
```

Hairball. Keep forgetting there are 32 teams in the NFL.

```{r}
clusterData <- dataMerged %>%
  select(TEAM,POS,PPR,YEAR) %>%
  group_by(TEAM,POS,YEAR) %>%
  mutate(TEAM_RANK = row_number()) %>%
  filter(TEAM_RANK<=2 & TEAM!='3TM' & TEAM!='4TM') %>%
  arrange(TEAM,POS,YEAR,PPR) %>% 
  filter(TEAM!='2TM') %>%
  unite(POS_TEAM_RANK,POS,TEAM_RANK,sep = '_') %>%
  unite(TEAM_YEAR,TEAM,YEAR,sep = '_') %>%
  select(TEAM_YEAR,POS_TEAM_RANK,PPR) %>%
  filter(!POS_TEAM_RANK %in% c('QB_2','NA_1','NA_2','RB_2','TE_2')) %>%
  spread(POS_TEAM_RANK,PPR)
clusterData
```

```{r}
cor(clusterData %>% select(-one_of('TEAM_YEAR')))
```

Takeways from correlations:

* QB and Pass Catcher relationship is huge, which is obvious. The better the QB
the better the other players and vice versa.

```{r}
write_csv(clusterData,'historicalFantasyTeamsClusterData.csv')
```

